---
title: "MY-DATA607-Week04-Chess"
author: "Michael Y."
date: "September 23, 2018"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    keep_md: yes
    theme: cerulean
  pdf_document: default
---

## Week 4 - Project 1

### Parse a provided table of results from a Chess tournament, and perform the requested transformations

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(encoding = 'UTF-8')
```

##### Load libraries
```{r load-libraries, eval=T}
library(readr)
library(stringr)
library(R.utils)
```

##### Load the raw datafile
```{r load-raw-datafile, eval=T}
setwd("C:/Users/Michael/Dropbox/priv/CUNY/MSDS/201809-Fall/DATA607_Andy_Sabrina/Week04-Project 1")
inputfile <- "tournamentinfo.txt"
rawchess <- read_lines(inputfile)
head(rawchess)
```
#### The datafile contains `r length(rawchess)` lines.    
##### One of every three lines is a separator composed of hyphens, which we will ignore, while the subsequent pair of lines contains information about each participant and the opponents he/she faced in the tournament.
#### The information about the 64 players is preceded by a pair of lines with header titles.   

#### Here we will transform the above into an array of 65 lines (one for the header titles, followed by one for each of 64 players): 
```{r reshape-raw-input, eval=T}
TempOutputArray <- NULL
TempOutputLine <- ""
TempLineNum <- 0
for (row in 1:length(rawchess)) {
  if (row %% 3 == 1) {
    ### this row is just hyphens, so we will ignore it
    invisible(NULL)   #no-op
  }
  if (row %% 3 == 2) {
    ### This is the first of two lines for the player
    TempLineNum <- TempLineNum + 1
    TempOutputLine <- rawchess[row]
  }
  if (row %% 3 == 0) {
    ### This is the second of two lines for the player
    TempOutputLine <- str_trim(paste(TempOutputLine, rawchess[row]))
    TempOutputArray[TempLineNum] <-TempOutputLine
  }
}
head(TempOutputArray)
```

#### Split at the pipe separators into elements, and trim the extra spaces preceding and following each element:
```{r split-and-trim, eval=T}
TrimmedOutputArray <- lapply(strsplit(x=TempOutputArray, split = "\\|"),str_trim)
head(TrimmedOutputArray)
```


#### The column titles are not unique, and do not always make sense:  

```{r bad-column-titles, eval=T}
BadColNames <- TrimmedOutputArray[[1]]
BadColNames
```

#### so we will replace them with better titles:  
```{r improve-column-titles, eval=T}
BetterColNames <- c("ID", 
                    "PlayerName", 
                    "TotalPts", 
                    sprintf("Round%d",1:7), 
                    "State", 
                    "USCFID_RatePre_RatePost", 
                    "Pts", 
                    sprintf("Color%d",1:7))
BetterColNames
TrimmedOutputArray[[1]] = BetterColNames

```

#### Now convert the above array into a **matrix**, where the first line from the array becomes the column names, and the remaining 64 lines become the data:

```{r make-array-into-matrix, eval=T}
TotalRows <- length(TrimmedOutputArray)
TotalElements <- length(unlist(TrimmedOutputArray))
ColumnCount <- table(unlist(lapply(TrimmedOutputArray,length)))
TrimmedOutputMatrix  <-  matrix(data = unlist(TrimmedOutputArray[2:TotalRows]), 
                                nrow = length(TrimmedOutputArray)-1, 
                                ncol=length(TrimmedOutputArray[[1]]), 
                                byrow = T)
colnames(TrimmedOutputMatrix) <- TrimmedOutputArray[[1]]
head(TrimmedOutputMatrix)
tail(TrimmedOutputMatrix)
```

#### There were `r TotalRows` total rows in the array, corresponding to `r TotalElements` total elements.   
##### This is consistent with having `r names(ColumnCount)` columns on each row.   

##### Ignore whatever number comes after a P, i.e., if a rating is nnnnPxx , use only the first part.

```{r extract-player-ratings}
df = data.frame(TrimmedOutputMatrix, stringsAsFactors = F)
USCFID <- as.integer(str_extract(df$USCFID_RatePre_RatePost, "^\\d{8}"))
df$USCFID <- USCFID
tempPrerating1 <- str_replace(string = df$USCFID_RatePre_RatePost, "\\d{8} / R: *","")
tempPrerating2 <- str_replace(tempPrerating1, "->.*$","")
tempPrerating3 <- str_replace(tempPrerating2,"P.*$","")
df$RatePre <- as.integer(tempPrerating3)
cat('Pre-tournament ratings for each player:\n')
df$RatePre

# we're not actually using the Post-tournament ratings, but they might be useful to have...
tempPostrating1=str_trim(str_replace(string = df$USCFID_RatePre_RatePost, "^.*->",""))
tempPostrating2=str_replace(tempPostrating1, "P.*$","")
df$RatePost <- as.integer(tempPostrating2)
cat('Post-tournament ratings for each player:\n')
df$RatePost
```

#### separate out the win/loss/draw/etc. for each player and tally up, in order to determine now many games each played:
```{r separate-win-loss-draw, eval=T}
### WLD is a 64x7 grid of the individual letters correspondig to each game played (or, not played) per player/round
WLD = mapply(function(x) str_extract(string = x, pattern = "^."),subset(x = df, select = Round1:Round7))
cat('Win-Loss-Draw for the various players: \n')
tail(WLD)

### collapseWLD collapses each row into a single string
collapseWLD = apply(X=WLD,MARGIN = 1, function(rw) str_c(rw,collapse=''))
cat('The above, collapsed into a string for each player: \n')
tail(collapseWLD)
```

#### Determine how many games were W, L, or D (ignoring other items): 
```{r table-WLD, eval=T}
### compute a table for each row, where the table counts the occurance of each value (W, L, D, or other indicators)
tempResults1 <- apply(X=WLD, MARGIN = 1, FUN=table)
cat('table tallying results for each competitor (only the final 6 displayed here:) \n')
tail(tempResults1)
```


#### tally the number which are "W", "L", or "D" -- (other values such as "U" or "B" indicate that no game was played):
```{r tally-table-totals, eval=T}
numgames <- lapply(X = tempResults1, 
                   FUN = function(rw) ifelse(is.na(rw["W"]),0,rw["W"])
                                     +ifelse(is.na(rw["L"]),0,rw["L"])
                                     +ifelse(is.na(rw["D"]),0,rw["D"]))
df$numgames <- as.integer(numgames)
cat('Number of "W", "L", "D" games played by each player: \n')
df$numgames
```

####Extract the list of opponents from the Round1:Round7 columns, so we can look up their ratings:     
######(Note that there are "NA" values in cases where no game was played)
```{r extract-opponents, eval=T}
opponents=mapply(function(x) as.integer(str_trim(str_replace(string = x, pattern = "^.",""))),
                 subset(x = df, select = Round1:Round7))
colnames(opponents) <- sprintf("Opp%d",1:7)
opponents

```

##### Append the 7 opponent columns to the dataframe: 

```{r append-opponent-ID-columns-to-df, eval=T}
### CAUTION -- this would duplicate columns if this chunk is re-evaluated manually, 
### but this is not a problem when knitting the entire result
df <- cbind(df,opponents)
head(df)
```


#### Look up the ratings for each opponent:
```{r lookup-opponent-ratings, eval=T}
oppPreRatings=apply(opponents,
                    MARGIN = c(1,2), 
                    FUN=function(x)df[x,]$RatePre)
### set the the column names
colnames(oppPreRatings) <- sprintf("OppPreRate%d",1:7)
cat('Opponent ratings, prior to the tournament: \n')
```


#### Append the 7 opponent rating columns to the dataframe:

```{r append-opponent-rating-columns-to-df, eval=T}
### CAUTION -- this would duplicate columns if this chunk is re-evaluated manually, 
### but this is not a problem when knitting the entire result

oppPreRatings
df <- cbind(df,oppPreRatings)
```

##### Compute the average opponent rating, for each participant: 
```{r average-opponent-ratings, eval=T}
AvgOppRatePre <- rowSums(oppPreRatings,na.rm=T)/df$numgames
cat('Average opponent rating, prior to the tournament: \n')
AvgOppRatePre
### Append the result to the dataframe
df$AvgOppRatePre <- AvgOppRatePre
```

##### doublecheck to be sure that rowSums/numgames gives the same result as rowMeans: 

```{r doublecheck-rowMeans}
rowMeans(oppPreRatings,na.rm = T)
### any differences?
rowSums(oppPreRatings,na.rm=T)/df$numgames - rowMeans(oppPreRatings,na.rm = T)
```

##### The input data has the names entirely capitalized:   

```{r all-caps-names, eval=T}
cat('Initial name format supplied in input data: \n')
head(df$PlayerName)
```

##### but the sample output shows just the first letter on each name component capitalized:
```{r change-name-capitalization, eval=T}
##library(R.utils) ## loaded at top

# lowercase each name, and then separate each into an array of strings, so "capitalize"" will operate on each component
CapitalList <- lapply(strsplit(x = tolower(df$PlayerName), 
                              split = " "),
                     capitalize)
cat('Properly Capitalized name components (in list format): \n')
head(CapitalList)
```

#### paste the resulting names back together again
```{r paste-name-back-together, eval=T}

CapitalNameArray=unlist(lapply( X = CapitalList, 
                                FUN= function(name) paste(name, collapse=" ")))
cat('Each name pasted back together again, in array format: \n')
head(CapitalNameArray)

# add Capitalized Name to the data frame
df$CapitalizedName = CapitalNameArray
```

##### Select output columns for final result into smaller dataframe:
```{r select-output-columns, eval=T}
OutputColumns = c("CapitalizedName","State", "TotalPts", "RatePre", "AvgOppRatePre")
OutputDF <- subset(x = df, select = OutputColumns)

##### improve the column headers, to match the assignment
OutputHeaders <- c("Player's Name", 
                   "Player's State", 
                   "Total Number of Points", 
                   "Player's Pre-Rating",
                   "Average Pre Chess Rating of Opponents")
colnames(x = OutputDF) <- OutputHeaders
head(OutputDF)
```

#### write the results into a .csv file: 

```{r write-results-to-csv, eval=T}
head(OutputDF)
options(encoding = 'UTF-8')
filename <- "MY-DATA607-Week04-ChessData.csv"
write.csv(OutputDF, filename, row.names=FALSE)
```

#### re-read the results and display them:
```{r reread-results, eval=T}
re_read_output <- read_lines(filename)
cat (re_read_output, sep = '\n')
```

