---
title: "MY607-Week03-MichaelY"
author: "Michael Y."
date: "September 16, 2018"
output:
  pdf_document: default
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###### Load some packages
```{r}
library(stringr)
library(readr)

```
###### fix a directory
```{r}
setwd("C:/Users/Michael/Dropbox/priv/CUNY/MSDS/201809-Fall/DATA607_Andy_Sabrina/Week03/")
```


## Exercise 3.
###Copy the introductory example. The vector ***name*** stores the extracted names.
```{r}
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5543642Dr. Julius Hibbert"
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
phone <- unlist(str_extract_all(raw.data, "\\(?(\\d{3})?\\)?(-| )?\\d{3}(-| )?\\d{4}"))
phone
data.frame(name = name, phone = phone)

```


####Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.

The above description is a bit unclear -- I am unsure as to whether this means that titles should be stripped out.     
Also, in the case of an individual like Mr. Burns, are we to use just the letter "C" as his first name?    

I am going to make the assumption that the question is asking us to take those names which are presented as:    

######lastname, firstname    

and rearrange such names so that the comma is removed and the names are presented in the order:   

######firstname lastname    

but in the case of any titles, initials, etc., I am not going to remove them.    

(The instructions should have been more clear if something different was desired.)    

##### **STARTING NAMES:**
```{r, eval=T}
name
```

```{r, eval=T}

for (i in 1:length(name)) {
  if (str_detect(string = name[i], pattern = ",")) {
    origname <-unlist(str_split(string = name[i], pattern = ","))
    flippedname <- str_trim(paste(origname[2], origname[1]))
    name[i] <- flippedname
  }
}
```

##### **ENDING NAMES:**
```{r, eval=T}
name
```
###### Note the changes to the names in the center column. 
    
    
    
####Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).

```{r}
hastitle = str_detect(string = name, pattern="^Rev\\.|^Dr\\.")
hastitle
name[hastitle]
```


####Construct a logical vector indicating whether a character has a second name.

```{r, eval=T}
### How many parts does each name have?
numnameparts = unlist(lapply(X = str_split(name," "), FUN = length))
numnameparts
cat(paste(numnameparts, name), sep="\n")
```

If the individual ***has*** a title (i.e.,. "Rev." or "Dr.", then his name must have ***4*** parts to include a middle name.   
If the individual does ***not*** have a title, then his name must have ***3*** parts to include a middle name.

```{r, eval=T}

hasmiddlename = (hastitle & numnameparts == 4) | (!hastitle & numnameparts == 3)
hasmiddlename
name[hasmiddlename]
```


#######################################################################

##Exercise 4.	

###Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

###1.	[0-9]+\\\\$

This matches one or more digits, followed by a dollar-sign "$" character.
(Note: It does **not** represent digits at the end of a line, because the backslashes preceding the dollar-sign give it its literal meaning, rather than its "end-of-line" meaning.)

```{r regex, eval=T}
regex1 = '[0-9]+\\$'

a = '1$'
b = "2$"
c = '34$'
d = "56$"
e = 'xyz7$abc456$789'
f = "xyz89$abc$234$567"
g = 'xyz10$abc$asdf$'
h = "xyz11$abc$123456$asdfg"
x = "This is a long string with letters$9876 and 1234$numbers - many 56789$numbers"
example1 <- c(a,b,c,d,e,f,g,h,x)

unlist(str_extract_all(example1, regex(regex1)))
##example1[grep(regex1, x=example1)]
unlist(regmatches(x=example1, m=gregexpr(text=example1, pattern=regex1)))

```

##########################################################

###2.	\\\\b[a-z]{1,4}\\\\b

Thus matches any "word" containing between 1 and 4 lower-case letters.  Note that in addition to spaces, a "word boundary" can also be delimited by certain characters such as a dollar-sign ("$")
```{r, eval=T}
regex2 = "\\b[a-z]{1,4}\\b"
example2a = "This is a long sentence containing many words. Now$is$the$winter$of$our$discontent."
example2b = "This.is.a.long.sentence.containing.many.words. Now-is-the-winter-of-our-discontent."
example2c = "This:is:a:long:sentence:containing:many:words. Now_is_the_winter_of_our_discontent."
res1=unlist(str_extract_all(example2a, regex(regex2)))
cat("\n____example2a_____", paste(1:length(res1),res1), "__________________\n",sep='\n')

res2=unlist(str_extract_all(example2b, regex(regex2)))
cat("\n____example2b_____", paste(1:length(res2),res2), "__________________\n",sep='\n')

#### the underscores do not constitute word separators, so the last portion will not be captured
res3=unlist(str_extract_all(example2c, regex(regex2)))
cat("\n____example2c_____", paste(1:length(res3),res3), "__________________\n",sep='\n')

```
######Note that the underscores in example2c do not constitute word separators, thus none of the words are taken from the last part of that example.

##########################################################

###3.	.*?\\\\.txt$  

This matches any number of characters (of any kind) followed by ".txt" , which must then conclude the line.

```{r, eval=T}
regex3 = ".*?\\.txt$"
example3 = c(
  "This is a filename.txt", 
  "Hello123456.txt", 
  "what is new.txt",
  "This line will not be selected because .txt is not at the end.",
  "What if we put a linefeed following .txt\nAnd see what happens?"
)
result3=unlist(str_extract_all(example3, regex(regex3)))
result3
cat(result3, sep="\n")

```

##########################################################

###4.	\\\\d{2}/\\\\d{2}/\\\\d{4}

This matches the usual way of displaying a date, e.g., mm/dd/yyyy (in the USA) or dd/mm/yyyy (elsewhere.)

```{r, eval=T}
regex4 = "\\d{2}/\\d{2}/\\d{4}"
example4 = c(
"Today is 09/16/2018. Yesterday was 09/15/2018. But, Europeans write 15/09/2018.", 
"Of course, there are no controls restricting months to be (01:12) or days to be (01:31)",
"So we could write 00/00/0000 or 99/99/9999 or 12/34/5678 and it would still be selected.",
"However, we can only use digits, not numbers, so we will not get ab/cd/efgh ."
)
unlist(str_extract_all(example4, regex(regex4)))
```


##########################################################

###5.	<(.+?)>.+?</\\\\1>

This identifies an HTML or XML-style block where some tag references the start and end of such block, with ***non-empty*** text in between.  
The start tag would be something like \<FOO\> and the corresponding end tag would be \</FOO\>.
```{r, eval=T}
regex5 = "<(.+?)>.+?</\\1>"
example5 = c(
  "<HTML>some content</HTML>",
  "<b>text content inside block</b>",
  "<bad>this will not work<bad>",
  "</bad2>also this will not work<bad2>",
  "<empty></empty>",
  "<notempty>x</notempty>",
  "</maybe>perhaps this will work?<//maybe>"
)
unlist(str_extract_all(example5, regex(regex5)))


```

##########################################################

## Exercise 9  

####The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others! The code snippet is also available in the materials at www.r-datacollection.com.
```{r, eval=T}
### The code snippet with the secret message is here:
### http://www.r-datacollection.com/materials/regex/code_exercise.txt

#library(readr)        # loaded above

fileURL="http://www.r-datacollection.com/materials/regex/code_exercise.txt"
secret <- read_file(fileURL)
secret
```
Let's see if any of the regex classes can help filter out something meaningful:
```{r, eval=T}
regexclasses <- c(
"[[:digit:]]", 	#Digits: 0 1 2 3 4 5 6 7 8 9
"[[:lower:]]", 	#Lowercase characters: a-z
"[[:upper:]]", 	#Uppercase characters: A-Z
"[[:alpha:]]", 	#Alphabetic characters: a-z and A-Z
"[[:alnum:]]", 	#Digits and alphabetic characters
"[[:punct:]]", 	#Punctuation characters: . , ; etc.
"[[:graph:]]", 	#Graphical characters: [:alnum:] and [:punct:]
"[[:blank:]]", 	#Blank characters: Space and tab
"[[:space:]]", 	#Space characters: Space, tab, newline, and other space characters
"[[:print:]]"   #Printable characters: [:alnum:], [:punct:] and [:space:]
)

regexclasses
```

Let's try each regex class on the secret, and see if any of them gives insight:
```{r, eval=T}
resultlist = lapply(X=str_extract_all(string = secret, pattern=regexclasses), 
                    FUN=str_c, collapse="")
unlist(resultlist)
```

The third one shows promise -- it has some discernable words, but they are run together.

```{r, eval=T}
regexclasses[3]
resultlist[[3]]
```

I see some punctuation in the sixth item:

```{r, eval=T}
regexclasses[6]
resultlist[[6]]
```

Perhaps we should try combining both regexes (i.e., 3 and 6) ?

```{r, eval=T}
combinedregex=sub(pattern="\\]\\[", replacement="", x = paste0(regexclasses[3],regexclasses[6]))
combinedregex
```

Here's the result from combining the UPPERCASE and the punctuation:
```{r, eval=T}
result9=str_c(unlist(str_extract_all(string = secret, pattern=combinedregex)), collapse="")
result9
```

OK, time to clean up the punctuation.  Since the original secret string didn't contain any spaces, clearly it used a dot in place of a space.  So, let's switch them back:
```{r, eval=T}
result9a = str_replace_all(result9,"\\."," ")
result9a
```

To make this grammatically correct, it needs a comma:
```{r, eval=T}
result9b = str_replace(result9a," ",", ")
result9b
```

### The secret message is: "`r result9b`"

#### (Please, tell me something I ***didn't*** already know?)