---
title: "DATA607-Week1-Mushrooms"
author: "Michael Y."
date: "September 1, 2019"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    code_folding: show
    toc: no
    toc_float: no
    toc_depth: 3
  pdf_document:
    toc: no
classoption: landscape
editor_options:
  chunk_output_type: inline
---
<style>
  .main-container {
    max-width: 1200px !important;
  }
</style>

# Assignment 1 - Loading Data into a Data Frame - Mushroom Dataset

```{r setup, eval=T}
knitr::opts_chunk$set(echo = TRUE)
directory = "C:/Users/Michael/Dropbox/priv/CUNY/MSDS/201909-Fall/DATA607_Tati_Andy/20190901_Week01/"
knitr::opts_knit$set(root.dir = directory)

### Make the output wide enough
options(scipen = 999, digits=6, width=120)

### Load some libraries
library(tidyr)
library(dplyr)
library(kableExtra)

```


## Mushrooms Dataset
A famous-if slightly moldy-dataset about mushrooms can be found in the UCI repository here: 
https://archive.ics.uci.edu/ml/datasets/Mushroom
The fact that this is such a well-known dataset in the data science community makes it a good dataset to use for comparative benchmarking.  For example, if someone was working to build a better decision tree algorithm 
(or other predictive classifier) to analyze categorical data, this dataset could be useful.  A typical problem (which is beyond the scope of this assignment!) is to answer the question, 

### "Which other attribute or attributes are the best predictors of whether a particular mushroom is poisonous or edible?"

Your task is to study the dataset and the associated description of the data (i.e. "data dictionary"). You may need to look around a bit, but it's there! 
You should take the data, 


```{r obtain-the-data-from-UCI, eval=T}
### Avoid setwd when knitting -- instead use above knitr::opts_knit$set(root.dir = directory)
#setwd("C:/Users/Michael/Dropbox/priv/CUNY/MSDS/201909-Fall/DATA607_Tati_Andy/20190901_Week01/")
download.file('https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data', 'mushroom-database.csv')
```
### Create a data frame with a subset of the columns in the dataset. 
(First, I'll create a dataframe with all 23 columns:)
```{r make_dataframe, eval=T}
mushroom_df <- read.csv('mushroom-database.csv', header=FALSE, stringsAsFactors=TRUE)
```
According to the documentation, the size of the data frame should be 8124x3. 
Checking:
```{r check-df-dimension, eval=T}
dim(mushroom_df)

```

### Determine which lines represent  poisonous mushrooms by converting to a 0/1 variable, usable in logistic regression 
```{r poisonous, eval=T}
poisonous <- mushroom_df$V1=="p"
poisonous=as.integer(poisonous)
head(mushroom_df$V1)
head(poisonous)
```

### Get metadata -- info about the data

```{r get-metadata, eval=T}
download.file('https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.names', 'mushroom-info.txt')
# read the lines into R
mushroom_meta=readLines(con="mushroom-info.txt")
```

### We only care about metalines 106-140

```{r trim-metadata, eval=T}
mushroom_meta2 = mushroom_meta[106:140]   
# trim the whitespace on the overflow lines
mushroom_meta2 = trimws(mushroom_meta2)
kable(mushroom_meta2, caption="Mushroom Metadata text") %>%
  kable_styling(c("striped", "bordered"))
```

### Attributes with many values are split onto two lines
Lines starting an attribute start with a number, while rollover lines start with a letter.
Let's create a function which will join the rollover lines onto the starting lines:

```{r pastearray-function, eval=T}

pastearray = function(chararray) {
# take an array of character strings
# some lines start with a number, while other lines start with a letter
# the lines which start with a letter are to be pasted at the end of the preceding line, 
# and the array is to be shortened
  
# first, determine which lines start with a number, and which with a letter
  # which lines in chararray start with a number, vs a letter?
  firstc <- substr(chararray,start = 1,stop = 1)          #Select the first character from each line
  firstc_numeric <- suppressWarnings(as.numeric(firstc))  # cast each character as numeric; characters return NA
                                                          # I don't want to see the warning messages, so suppress them
  firstc_not_numeric = sapply(firstc_numeric,is.na)       # true for each continuation line
  firstc_is_numeric = !firstc_not_numeric
  firstc_is_numeric
  chararray[firstc_is_numeric]
  
  tempoutputarray = NULL
  tempoutputline = ""
  j = 0
  for (i in 1:length(chararray))
    {  
    if (firstc_is_numeric[i]) {
      # we are starting a new line, so print out the prior line assemblage -- 
      if (i>1) {
      # unless we are at the very beginning, in which case there is nothing to print      
        j=j+1
        tempoutputarray[j]=tempoutputline
        #####print(paste(j,": ",  tempoutputarray[j]))
        # reset to blank line
        tempoutputline=""
      }  
    # set this line as the newline
    tempoutputline = chararray[i]  
    }
    else if (firstc_not_numeric[i])  # we are on a continuation !!!!! 
      {
      tempoutputline = paste0(tempoutputline, chararray[i])   # paste this line onto the previous line
      #####print(paste("*** pasting ", i, j, tempoutputline))
    }
  }
  # when we reach the end, we still have to print out the prior line
  j=j+1
  tempoutputarray[j]=tempoutputline
  #####print(paste(j,": ",  tempoutputarray[j]))
  #####print ( paste(i,",",chararray[i]))
  return(tempoutputarray)
}
```

Apply this function to the above mushroom_meta 

```{r run-pastearray, eval=T}
mushroom_meta2 = pastearray(mushroom_meta2)
mushroom_meta2
```

Success, now there are only 23 rows, each representing one column.

Modify the first line to resemble the other lines. 
Since ascertaining whether poisonous vs. edible is the TARGET, I'll relabel this line as such.
```{r replace-first-line, eval=T}
############
############
firstline = mushroom_meta2[1]
######firstline
# change the front part of the line
firstline = sub(pattern="7. Attribute Information: (classes:", replacement="0. TARGET:                  ", x=firstline, fixed=TRUE)
######firstline
# change the rear part of the line - remove the space and the final right parens
firstline = sub(pattern=", poisonous=p)", replacement=",poisonous=p", x=firstline, fixed=TRUE)
######firstline
#################################
mushroom_meta2[1]=firstline
mushroom_meta2
###################################

```

Extract the list of attribute names from the above
```{r get_attribute_names, eval=T}
names1=gsub(pattern=":.*$", replacement = "", x=mushroom_meta2)
#names1
names2=gsub(pattern="^[0-9]*. ", replacement = "", x=names1, perl=FALSE)
#names2
# replace hyphens with underscores or it will cause problems later
names3=gsub(pattern="-", replacement="_", x=names2)
kable(names3, caption="Mushroom Attribute Names") %>%
  kable_styling(c("striped", "bordered"))
```

extract the list of factors (their descriptive names, and their single character abbreviations) from each line
```{r get_factors, eval=T}
factors1 = gsub(pattern="^.*:  *",replacement="",x=mushroom_meta2)
factors1
```
### Rename the factors with the descriptive names
(I had to set this up manually because I couldn't get the proper processing of the quotation marks in R)
```{r rename-the-factors, eval=T}

levels(mushroom_df$V1)
levels(mushroom_df$V1) <-  list(edible="e",poisonous="p")
levels(mushroom_df$V1)
levels(mushroom_df$V2) <-  list(bell="b",conical="c",convex="x",flat="f",knobbed="k",sunken="s")
levels(mushroom_df$V3) <-  list(fibrous="f",grooves="g",scaly="y",smooth="s")
levels(mushroom_df$V4) <-  list(brown="n",buff="b",cinnamon="c",gray="g",green="r",pink="p",purple="u",red="e",white="w",yellow="y")
levels(mushroom_df$V5) <-  list(bruises="t",no="f")

levels(mushroom_df$V6)
levels(mushroom_df$V6) <-  list(almond="a",anise="l",creosote="c",fishy="y",foul="f",musty="m",none="n",pungent="p",spicy="s")
levels(mushroom_df$V6)

levels(mushroom_df$V7) <-  list(attached="a",descending="d",free="f",notched="n")
levels(mushroom_df$V8) <-  list(close="c",crowded="w",distant="d")
levels(mushroom_df$V9) <-  list(broad="b",narrow="n")
levels(mushroom_df$V10) <-  list(black="k",brown="n",buff="b",chocolate="h",gray="g",green="r",orange="o",pink="p",purple="u",red="e",white="w",yellow="y")
levels(mushroom_df$V11) <-  list(enlarging="e",tapering="t")
levels(mushroom_df$V12) <-  list(bulbous="b",club="c",cup="u",equal="e",rhizomorphs="z",rooted="r",missing="?")
levels(mushroom_df$V13) <-  list(fibrous="f",scaly="y",silky="k",smooth="s")
levels(mushroom_df$V14) <-  list(fibrous="f",scaly="y",silky="k",smooth="s")
levels(mushroom_df$V15) <-  list(brown="n",buff="b",cinnamon="c",gray="g",orange="o",pink="p",red="e",white="w",yellow="y")
levels(mushroom_df$V16) <-  list(brown="n",buff="b",cinnamon="c",gray="g",orange="o",pink="p",red="e",white="w",yellow="y")
levels(mushroom_df$V17) <-  list(partial="p",universal="u")
levels(mushroom_df$V18) <-  list(brown="n",orange="o",white="w",yellow="y")
levels(mushroom_df$V19) <-  list(none="n",one="o",two="t")
levels(mushroom_df$V20) <-  list(cobwebby="c",evanescent="e",flaring="f",large="l",none="n",pendant="p",sheathing="s",zone="z")
levels(mushroom_df$V21) <-  list(black="k",brown="n",buff="b",chocolate="h",green="r",orange="o",purple="u",white="w",yellow="y")
levels(mushroom_df$V22) <-  list(abundant="a",clustered="c",numerous="n",scattered="s",several="v",solitary="y")
levels(mushroom_df$V23) <-  list(grasses="g",leaves="l",meadows="m",paths="p",urban="u",waste="w",woods="d")
```

### Display mushroom df summary
```{r summary_mushroom_df,eval=T}
summary(mushroom_df)
```
### Now, replace the names of the columns in the data set with descriptive attribute names (with underscores replacing hyphens:
```{r rename-the-columns, eval=T}
names(mushroom_df)=names3
head(mushroom_df)
```


```{r summary2, eval=T}
summary(mushroom_df)

```

### According to the documentation, a small number of columns provide an excellent prediction of which mushrooms are poisonous:

#### These columns are odor, spore-print-color, stalk-surface-below-ring, and stalk-color-above-ring 




#### Rule #1: odor (V6) is not Almond ("a"), Anise ("l"), or None ("n")

```{r rules, eval=T}

# Rule1: odor (V6) is NOT (Almond , Anise , or None) 
rule1 = (!(mushroom_df$odor == "almond" | mushroom_df$odor == "anise" | mushroom_df$odor == "none"))
sum(as.integer(rule1))
```

#### Rule #2: spore-print-color (V21) is Green ("r")
```{r rule2, eval=T}
# Rule2: spore-print-color is Green
rule2 = (mushroom_df$spore_print_color== "green")
sum(as.integer(rule2))

```

#### Rule #3: odor (V6) is None ("n") AND stalk-surface-below-ring (V14) is Scaly ("y") AND stalk-color-above-ring (V15) is NOT brown (n)

```{r rule3, eval=T}
# Rule3: odor is None AND 
#       stalk-surface-below-ring is Scaly AND 
#       stalk-color-above-ring is NOT brown
rule3 = ((mushroom_df$odor == "none") & (mushroom_df$stalk_surface_below_ring == "scaly" )&( mushroom_df$stalk_color_above_ring != "brown"))
sum(as.integer(rule3))
```

### run logistic regression using just rule 1
```{r logistic1, eval=T}
mmodel1 <- glm(poisonous ~ rule1, data = mushroom_df, family="binomial")
summary(mmodel1)
model1predictor <- (as.integer(mmodel1$residuals)==1)
model1errors <- sum(as.integer(model1predictor != poisonous))
model1errors
```

### run logistic regression using rules 1 and 2
```{r logistic12, eval=T}
mmodel2 <- glm(poisonous ~ rule1 + rule2, data = mushroom_df, family="binomial")
summary(mmodel2)
model2predictor <- (as.integer(mmodel2$residuals)==1)
model2errors <- sum(as.integer(model2predictor != poisonous))
model2errors
```

### run logistic regression using rules 1, 2 and 3
```{r logistic123,eval=T}
mmodel3 <- glm(poisonous ~ rule1 + rule2 + rule3, data = mushroom_df, family="binomial")
summary(mmodel3)
model3predictor <- (as.integer(mmodel3$residuals)==1)
model3errors <- sum(as.integer(model3predictor != poisonous))
model3errors
```

Combining the first three rules gives a result which misses only 8 poisonous mushrooms.

### Subsetting the data

You should include the column that indicates edible or poisonous and three or four other columns. 

You should also add meaningful column names and replace the abbreviations used in the data-for example, in the appropriate column, "e" might become "edible." 

```{r subsets, eval=T}
mushroom_subsetdf = subset(mushroom_df, select=c(TARGET, odor, spore_print_color, stalk_surface_below_ring, stalk_color_above_ring))
head(mushroom_subsetdf)

```

### Summary of subsets
```{r subsetsummary, eval=T}
summary(mushroom_subsetdf)
```

### Your deliverable is the R code to perform these transformation tasks.

