---
title: "MY-DATA607-Week09-NYTimes-API"
author: "Michael Y."
date: "October 27, 2019"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
    keep_md: yes
  html_document:
    highlight: pygments
    theme: cerulean
    code_folding: show
    toc: yes
    toc_float: yes
    toc_depth: 3
classoption: landscape
editor_options:
  chunk_output_type: inline
---
<style>
  .main-container {
    max-width: 1200px !important;
  }
</style>
---

# Week 9 - Assignment - Web APIs (NY Times)

The New York Times web site provides a rich set of APIs, as described [here](http://developer.nytimes.com/docs).

You’ll need to start by signing up for an API key.

Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/Michael/Dropbox/priv/CUNY/MSDS/201909-Fall/DATA607_Tati_Andy/20191027_Week09")
```


#### Load libraries
```{r load-libraries, message=FALSE, warning=FALSE}
library(XML)
library(rlist)
library(knitr)
library(kableExtra)
library(jsonlite)
library(RCurl)
library(tidyr)
library(dplyr)
```



### Since Hallowe'en is in a few days, let's look for articles on this topic
```{r retrieve_NYT, eval=T}

# API key from website registration
MY_api_key <- "&api-key=WUNJOM1fVGLXpEzPYnHXnhONs8fFYU9V"

# search for articles with keyword "halloween"
url <- "http://api.nytimes.com/svc/search/v2/articlesearch.json?q=halloween"
request <- paste0(url, "&", MY_api_key)
result <- fromJSON(request)

```

```{r process_results, eval=T}

#pull $docs from result
NYT_docs <- result$response$docs

### Confirm that it's a dataframe
class(NYT_docs)

#get the column names
NYT_colnames <- colnames(NYT_docs)
t(t(NYT_colnames))
```

There is a lot of data here, not all of it interesting...

### pull web_url from NYT_docs
```{r web_url, eval=T, warning=F}
NYT_web_url <- NYT_docs$web_url
NYT_web_url %>% 
  kable(col.names="Web_URL",row.names=1:10) %>%  
  kable_styling(c("striped", "bordered"))
```

### pull snippet from NYT_docs
```{r snippet, eval=T, warning=F}
NYT_snippet <- NYT_docs$snippet
NYT_snippet %>% 
  kable(col.names="Snippet",row.names=1:10) %>%  
  kable_styling(c("striped", "bordered"))
```


### pull lead_paragraph from NYT_docs
```{r lead paragraph, eval=T, warning=F}
NYT_lead <- NYT_docs$lead_paragraph
NYT_lead %>% 
  kable(col.names="Lead Paragraph",row.names=1:10) %>%  
  kable_styling(c("striped", "bordered"))
```

### pull abstract from NYT_docs
```{r abstract, eval=T, warning=F}
NYT_abstract <- NYT_docs$abstract
NYT_abstract %>% 
  kable(col.names="Abstract",row.names=1:10) %>%  
  kable_styling(c("striped", "bordered"))
```

### fashion my own data frame
```{r MY_NYT_df, eval=T, warning=F}
MY_NYT_df <- data.frame(NYT_web_url, NYT_snippet, NYT_lead, NYT_abstract)
MY_NYT_df %>% kable(row.names=1:10) %>% 
  column_spec(.,column = 1, width = "10em") %>% 
  kable_styling(c("striped", "bordered"))
```
