---
title: "MY607-Week04-Project1-Chess"
author: "Michael Y."
date: "September 22, 2019"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    code_folding: show
    toc: yes
    toc_float: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: 3
classoption: landscape
editor_options:
  chunk_output_type: inline
---
<style>
  .main-container {
    max-width: 1200px !important;
  }
</style>---

# Week 4 - Project 1

## Parse a provided table of results from a Chess tournament, and perform the requested transformations

### Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
directory = "C:/Users/Michael/Dropbox/priv/CUNY/MSDS/201909-Fall/DATA607_Tati_Andy/20190922_Week04/"
knitr::opts_knit$set(root.dir = directory)
options(encoding = 'UTF-8')
### Make the output wide enough
options(scipen = 999, digits=6, width=95)
```

#### Load libraries
```{r load-libraries, include=F}
library(readr)
library(stringr)
library(R.utils)
library(tidyr)
library(dplyr)
library(kableExtra)
```

\newpage
### Load the raw datafile (pull from github)
```{r load-raw-datafile, eval=T}
#Avoid using setwd for knitting - instead set directory above in knitr$opts_set
#setwd("C:/Users/Michael/Dropbox/priv/CUNY/MSDS/201909-Spring/DATA607_Tati_Andy_Sabrina/20190922_Week04")
### inputfile <- "tournamentinfo.txt"

### Obtain the file from github, rather than local drive
inputfile <- "https://raw.githubusercontent.com/myampol/MY607/master/tournamentinfo.txt"
rawchess <- read_lines(inputfile)
head(rawchess,18)
```

\newpage
### Analyse the raw datafile:
The datafile contains `r length(rawchess)` lines.       
One of every three lines is a separator composed of hyphens, which we will ignore, while the subsequent pair of lines contains information about each participant and the opponents he/she faced in the tournament.    
The information about the 64 players is preceded by a pair of lines with header titles.     
Here we will transform the above into an array of 65 lines (one for the header titles, followed by one for each of 64 players).    
### Transform the input data from `r length(rawchess)` lines into 65 lines:

```{r reshape-raw-input, eval=T}
TempOutputArray <- NULL
TempOutputLine <- ""
TempLineNum <- 0
for (row in 1:length(rawchess)) {
  if (row %% 3 == 1) {
    ### this row is just hyphens, so we will ignore it
    invisible(NULL)   #no-op
  }
  if (row %% 3 == 2) {
    ### This is the first of two lines for the player
    TempLineNum <- TempLineNum + 1
    TempOutputLine <- rawchess[row]
  }
  if (row %% 3 == 0) {
    ### This is the second of two lines for the player
    TempOutputLine <- str_trim(paste(TempOutputLine, rawchess[row]))
    TempOutputArray[TempLineNum] <-TempOutputLine
  }
}
head(TempOutputArray) %>% kable() %>%  kable_styling(c("striped", "bordered"))
```
\newpage
#### Split at the pipe separators into elements, and trim the extra spaces preceding and following each element:
```{r split-and-trim, eval=T}
# This is actually a list, but the column names are in the first element
TrimmedOutputArray <- lapply(strsplit(x=TempOutputArray, split = "\\|"),str_trim)

# extract the initial column names from the first element of the list
ColumnNames <- as.array(TrimmedOutputArray[[1]])

# create a "list" which has just the remaining 64 entries (sans column names)
TrimmedOutputList <- TrimmedOutputArray[-c(1)]

# make a data frame from the list (initially no column names)
TrimmedOutputDF <- as.data.frame(TrimmedOutputList,  cut.names = TRUE, stringsAsFactors = FALSE)

# this didn't quite give what we wanted, because each player is in a column 
# (i.e., there are 20 rows of features, and 64 columns of players)
dim(TrimmedOutputDF)

# show just the first two players (otherwise, too wide...)
head(TrimmedOutputDF[,1:2],14)  %>% kable() %>%  kable_styling(c("striped", "bordered"))

# so we need to transpose the data frame:
TrimmedOutputDF <- t(TrimmedOutputDF)
head(TrimmedOutputDF,3)

# however, the rownames and colnames are not as expected
# so, delete the row names
rownames(TrimmedOutputDF) <- NULL
# and, set the column names to the items extracted above
colnames(TrimmedOutputDF)<- ColumnNames
```

\newpage
```{r display-split-and-trim, eval=T}
# Display the info on the first 3 players (display transposed, for clarity)
t(head(TrimmedOutputDF,3))  %>% kable() %>%  kable_styling(c("striped", "bordered"))

```

\newpage
### Clean up column headers
#### The column titles are not unique, and do not always make sense:  

```{r bad-column-titles, eval=T}
BadColNames <- TrimmedOutputArray[[1]]
BadColNames 
```

#### so we will replace them with better titles:  
```{r improve-column-titles, eval=T}
BetterColNames <- c("ID", 
                    "PlayerName", 
                    "TotalPts", 
                    sprintf("Round%d",1:7), 
                    "State", 
                    "USCFID_RatePre_RatePost", 
                    "Pts", 
                    sprintf("Color%d",1:7))
BetterColNames
TrimmedOutputArray[[1]] = BetterColNames

```

\newpage
#### Now convert the above array into a **matrix**, where the first line from the array becomes the column names, and the remaining 64 lines become the data:

```{r make-array-into-matrix, eval=T}
TotalRows <- length(TrimmedOutputArray)
TotalElements <- length(unlist(TrimmedOutputArray))
ColumnCount <- table(unlist(lapply(TrimmedOutputArray,length)))
TrimmedOutputMatrix  <-  matrix(data = unlist(TrimmedOutputArray[2:TotalRows]), 
                                nrow = length(TrimmedOutputArray)-1, 
                                ncol=length(TrimmedOutputArray[[1]]), 
                                byrow = T)
colnames(TrimmedOutputMatrix) <- BetterColNames
head(TrimmedOutputMatrix) 
```

#### There were `r TotalRows` total rows in the array, corresponding to `r TotalElements` total elements.   
##### This is consistent with having `r names(ColumnCount)` columns on each row.   

\newpage

##### On the rating, ignore whatever number comes after a P, i.e., if a rating is nnnnPxx , use only the first part.

```{r extract-player-ratings}
df = data.frame(TrimmedOutputMatrix, stringsAsFactors = F)
USCFID <- as.integer(str_extract(df$USCFID_RatePre_RatePost, "^\\d{8}"))
df$USCFID <- USCFID
tempPrerating1 <- str_replace(string = df$USCFID_RatePre_RatePost, "\\d{8} / R: *","")
tempPrerating2 <- str_replace(tempPrerating1, "->.*$","")
tempPrerating3 <- str_replace(tempPrerating2,"P.*$","")
df$RatePre <- as.integer(tempPrerating3)
cat('Pre-tournament ratings for each player:\n')
df$RatePre

# we're not actually using the Post-tournament ratings, but they might be useful to have...
tempPostrating1=str_trim(str_replace(string = df$USCFID_RatePre_RatePost, "^.*->",""))
tempPostrating2=str_replace(tempPostrating1, "P.*$","")
df$RatePost <- as.integer(tempPostrating2)
cat('Post-tournament ratings for each player:\n')
df$RatePost
```

\newpage
### Compute wins/losses/draws and games played for each player

#### Separate out the win/loss/draw/etc. for each player and tally up, in order to determine now many games each played:
```{r separate-win-loss-draw, eval=T}
### WLD is a 64x7 grid of the individual letters corresponding to each game played 
### (or, not played) per player/round
WLD = mapply(function(x) str_extract(string = x, 
                                     pattern = "^."),
             subset(x = df, 
                    select = Round1:Round7))
cat('Win-Loss-Draw for the various players (first 6 displayed) : \n')
head(WLD)

### collapseWLD collapses each row into a single string
collapseWLD = apply(X=WLD,MARGIN = 1, function(rw) str_c(rw,collapse=''))
cat('The above, collapsed into a string for each player (first 6 displayed) : \n')
t(t(head(collapseWLD)))
```

#### Determine how many games were W, L, or D (ignoring other items): 
```{r table-WLD, eval=T}
### compute a table for each row, where the table counts the occurance of each value (W, L, D, or other indicators)
tempResults1 <- apply(X=WLD, MARGIN = 1, FUN=table)
cat('list of tables tallying results for each competitor (only the first 6 displayed here:) \n')
head(tempResults1)
```

\newpage
#### Tally the number which are "W", "L", or "D" -- (other values such as "U" or "B" indicate that no game was played):
```{r tally-table-totals, eval=T}
numgames <- lapply(X = tempResults1, 
                   FUN = function(rw) ifelse(is.na(rw["W"]),0,rw["W"])
                                     +ifelse(is.na(rw["L"]),0,rw["L"])
                                     +ifelse(is.na(rw["D"]),0,rw["D"]))
df$numgames <- as.integer(numgames)
cat('Number of "W", "L", "D" games played by each player: \n')
df$numgames
```

#### Extract the list of opponents from the Round1:Round7 columns, so we can look up their ratings:     
#### (Note that there are "NA" values in cases where no game was played)
```{r extract-opponents, eval=T}
opponents=mapply(function(x) as.integer(str_trim(str_replace(string = x, pattern = "^.",""))),
                 subset(x = df, select = Round1:Round7))
colnames(opponents) <- sprintf("Opp%d",1:7)
cbind(1:dim(opponents)[1],df$PlayerName,opponents) %>% 
  head() %>% kable() %>%  kable_styling(c("striped", "bordered"))

```

\newpage
### Determine the average ratings for each player's opponents, and append all relevant info to the dataframe:

#### Append the 7 opponent columns to the dataframe: 

```{r append-opponent-ID-columns-to-df, eval=T}
### CAUTION -- this would create duplicate columns 
### if this chunk (alone) were re-evaluated manually, 
### but this is not a problem when knitting the entire result, 
### as it can only be executed once
df <- cbind(df,opponents)
### First 3 players:
head(df,3)
```

\newpage
```{r display-transposed-results, eval-T}
### Transpose, for clarity:
t(head(df,3)) %>% kable() %>%  kable_styling(c("striped", "bordered"))
```

\newpage
#### Look up the ratings for each opponent:
```{r lookup-opponent-ratings, eval=T}
oppPreRatings=apply(opponents,
                    MARGIN = c(1,2), 
                    FUN=function(x)df[x,]$RatePre)
### set the the column names
colnames(oppPreRatings) <- sprintf("OppPreRate%d",1:7)
cat('Opponent ratings, prior to the tournament: \n')
head(cbind(1:dim(opponents)[1],df$PlayerName,oppPreRatings),22)  %>% 
  kable() %>%  kable_styling(c("striped", "bordered"))
```

\newpage
#### Append the 7 opponent rating columns to the dataframe:

```{r append-opponent-rating-columns-to-df, eval=T}
### CAUTION -- this would duplicate columns if this chunk is re-evaluated manually, 
### but this is not a problem when knitting the entire result
df <- cbind(df,oppPreRatings)
### display first 3 players, transposed, for clarity
t(head(df,n = 3))   %>% kable() %>%  kable_styling(c("striped", "bordered"))
```

\newpage
#### Compute the average opponent rating, for each participant: 
```{r average-opponent-ratings, eval=T}
# use na.rm=T to omit the NA values
# divide by the number of opponents played 
# (in the case where no game was played, opponent=NA and numgames < 7)
AvgOppRatePre <- rowSums(oppPreRatings,na.rm=T)/df$numgames
cat('Average opponent rating, prior to the tournament: \n')
AvgOppRatePre
### Append the result to the dataframe
df$AvgOppRatePre <- AvgOppRatePre
```

#### doublecheck to be sure that rowSums/numgames gives the same result as rowMeans: 

```{r doublecheck-rowMeans}
rowMeans(oppPreRatings,na.rm = T)
### any differences?
rowSums(oppPreRatings,na.rm=T)/df$numgames - rowMeans(oppPreRatings,na.rm = T)
```

\newpage
### Clean up the capitalization of player names:

#### The input data has the names entirely capitalized:   

```{r all-caps-names, eval=T}
cat('Initial name format supplied in input data: \n')
t(t(head(df$PlayerName)))
```

#### but the sample output shows just the first letter on each name component capitalized:
```{r change-name-capitalization, eval=T}
##library(R.utils) ## loaded at top

# lowercase each name, and then separate each into an array of strings, 
# so "capitalize"" will operate on each component
CapitalList <- lapply(strsplit(x = tolower(df$PlayerName), 
                              split = " "),
                     capitalize)
cat('Properly Capitalized name components (in list format): \n')
head(CapitalList)
```

#### paste the resulting names back together again
```{r paste-name-back-together, eval=T}

CapitalNameArray=unlist(lapply( X = CapitalList, 
                                FUN= function(name) paste(name, collapse=" ")))
cat('Each name pasted back together again: \n')
t(t(head(CapitalNameArray)))

# append Capitalized Name as a new column in the data frame
df$CapitalizedName = CapitalNameArray
```

\newpage
### Select output columns for final result into smaller dataframe:
```{r select-output-columns, eval=T}
OutputColumns = c("CapitalizedName","State", "TotalPts", "RatePre", "AvgOppRatePre")
FinalOutputDF <- subset(x = df, select = OutputColumns)

##### improve the column headers, to match the assignment
OutputHeaders <- c("Player's Name", 
                   "Player's State", 
                   "Total Number of Points", 
                   "Player's Pre-Rating",
                   "Average Pre Chess Rating of Opponents")
colnames(x = FinalOutputDF) <- OutputHeaders
head(FinalOutputDF,22) %>% kable() %>%  kable_styling(c("striped", "bordered"))
```

### Write the results into a .csv file: 

```{r write-results-to-csv, eval=T}
head(FinalOutputDF)
options(encoding = 'UTF-8')
results_filename <- "MY-DATA607-Week04-ChessData.csv"
write.csv(FinalOutputDF, results_filename, row.names=FALSE)
```

\newpage
### Re-read the results (stored in github, rather than local drive) and display them:
```{r reread-results, eval=T}
# read the results from local disk
# re_read_output <- read_lines(results_filename)

# Read the results from file (previously) posted to github
github_filename <- 
  "https://raw.githubusercontent.com/myampol/MY607/master/MY-DATA607-Week04-ChessData.csv"
re_read_output <- read_lines(github_filename)
cat (re_read_output, sep = '\n')
```

